---
title: "150_voi_test"
author: "Max Schofield"
date: "2022-10-21"
output:
  html_notebook:
  df_print: default
highlight: pygments
toc: yes
toc_float:
  toc_collapsed: true
toc_depth: 2

editor_options:
  chunk_output_type: inline
---

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Setup

Knitr options 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load R packages 

```{r load packages}
foo <- function(x){
  for( i in x ){
    #  require returns TRUE invisibly if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      #  Load package after installing
      require( i , character.only = TRUE )
    }
  }
}

#  Then try/install packages...
foo( c("tidyverse" , "bigrquery" ,"devtools", "DBI","glue", "lubridate", "here", "sf", "extrafont", "patchwork", "terra") )

# get fishwatch r independently 
if (!require("fishwatchr")) {devtools::install_github("GlobalFishingWatch/fishwatchr")}
library(fishwatchr)

```

Establish connection to BigQuery project

```{r con}
con <- DBI::dbConnect(drv = bigrquery::bigquery(), 
                      project = "world-fishing-827", 
                      use_legacy_sql = FALSE)
```

Set commonly used parameters for mapping

```{r}
# define bounding box/area of interest
# bbox <- data.frame(x_min = -8,
#                    x_max = 3,
#                    y_min = 0,
#                    y_max = 7)
# what projection should be used for mapping
best_proj <- fishwatchr::gfw_projections("Equal Earth")$proj_string
# best_proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```


# Establish Vessel of Interest 

Calling the table directly as it is saved due to being a large query

First question is establish how many vessels have the 150 MMSI prefix in total. 

```{r, echo=FALSE}
q_voi_ais <- c("SELECT * FROM `world-fishing-827.scratch_max.raw_AIS_150_mmsi_prefix_2021_20Oct2022`")

voi_ais <- fishwatchr::gfw_query(query = q_voi_ais,
                                 run_query = TRUE, 
                                 con = con)$data
n_distinct(voi_ais$ssvid)
```

This is querying the raw AIS data and there is `r n_distinct(voi_ais$ssvid)` vessels in total. 


```{r, echo=FALSE}
outside_chn <- readr::read_file(file = here::here("queries", ".", "outside_CHN.sql"))

outside_chn_vessels <- fishwatchr::gfw_query(query = outside_chn,
                                 run_query = TRUE,
                                 con = con)$data

n_distinct(outside_chn_vessels)

# restrict down to genuine VOI
outside_AIS <- filter(voi_ais, ssvid %in% c(outside_chn_vessels$ssvid))

# get rid of percentage - indicator of gear markers 
per <- filter(outside_AIS, stringr::str_detect(shipname, '\\%'))
head(table(per$shipname))

# get rid of ending in v - another indicator of gear markers
per_v <- filter(non_per, stringr::str_detect(shipname, 'v$|V$'))
table(per_v$shipname)

# get rid of ending in NET another indicator of gear markers
per_v_net <- filter(non_per_v, stringr::str_detect(shipname, 'NET'))
table(per_v_net$shipname)

# get rid of - in the name as another indicator of gear markers
per_v_net_dash <- filter(no_per_v_net, stringr::str_detect(shipname, '-'))
table(per_v_net_dash$shipname)

# remove vessels
no_per_v_net_dash <- outside_AIS %>% filter(!ssvid %in% unique(c(
    per$ssvid, 
    per_v$ssvid,
    per_v_net$ssvid, 
    per_v_net_dash$ssvid)))

# write out data to visaulise
write.csv(no_per_v_net_dash, './data/non_gear_outside_CHN.csv', row.names = F)

n_distinct(outside_AIS$ssvid)
n_distinct(no_per_v_net_dash$ssvid)

# need to be careful not to remove AIS which is 
voi <- unique(no_per_v_net_dash$ssvid)
df <- filter(voi_ais, ssvid %in% voi)

data <- no_per_v_net_dash[!is.na(no_per_v_net_dash$lat),]

```


```{r, echo=FALSE}

more_then10 <- data %>% group_by(ssvid) %>% summarise(n=n()) %>% filter(n>10)
n_distinct(more_then10$ssvid)

```

Of the `r n_distinct(voi_ais$ssvid)` total MMSIs associated with the 150 prefix,  `r n_distinct(outside_chn_vessels$ssvid)` transmitted outside the Chinese EEZ. 

Of the MMSIs operating outside China many have names which indicated they are associated 
with likely fishing gear. MMSIs associated with likely gear was removed from our vessel
of interest list based on removing MMSIs either transmitting a %, a Voltage indicator (V), 
the word 'NET' in the shipname, which are indicators of fishing gear. 

This filtering removed `r n_distinct(outside_AIS$ssvid) - n_distinct(no_per_v_net_dash$ssvid)`
MMSIs from our fleet of interest. Resulting in `r n_distinct(no_per_v_net_dash$ssvid)` remaining 
in our vessels of interest. 

Further filtering was conducted on vessels with less than 10 AIS transmissions between 2021 and 2022 which removed `r n_distinct(outside_AIS$ssvid) - n_distinct(more_then10$ssvid)` vessel from the analysis fleet leaving `n_distinct(more_then10$ssvid)` vessels. 

# Patterns in remaining vessel names  

The vessel of interest have five distinct vessel name groupings within them
1. names starting with FU of which there are `r n_distinct(filter(voi_ais, stringr::str_starts(shipname, 'FU'))$shipname)` distinct names with FU YUAN YU ... the most common pattern
2. names starting with HAI of which there are `r n_distinct(filter(voi_ais, stringr::str_starts(shipname, 'HAI'))$shipname)` distinct names with HAI YANG ... the most common pattern
3. names starting with LU RONG which there are `r n_distinct(filter(voi_ais, stringr::str_starts(shipname, 'LU'))$shipname)` distinct names with LURONG YUAN YU ... the most common pattern
4. names starting with SHUN HANG which there are `r n_distinct(filter(voi_ais, stringr::str_starts(shipname, 'SHUN'))$shipname)` distinct names following this pattern. This aligns with SHUN HANG which are an AIS unit manufacturer and may relate to units being reset.

One further vessel had the name LU QING YUAN YU 290. Only one MMSI transmitted this vessel name. 

Clear vessel names with some iterations in the dataset
- FU YUAN YU 715
- FU YUAN YU 717
- FU YUAN YU 9993
- FU YUAN YU 9994
- HAI HANG 1 
- HAI HANG 2
- HAI HANG 3
- HAI HANG 5
- HAI HANG 6 
- LU QING YUAN YU 290
- LU RONG YUAN YU 195 
- LU RONG YUAN YU 581 
- LU RONG YUAN YU 715 
- LU RONG YUAN YU 197 
- LU RONG YUAN YU 20 
- LU RONG YUAN YU 277 
- LU RONG YUAN YU 278 
- LU RONG YUAN YU 717 
- LU RONG YUAN YU 715 
- SHUN HANG 1 
- SHUN HANG 2
- SHUN HANG 3
- SHUN HANG 5
- SHUN HANG 6


# How many vessels have links to China through location   

Three MMSI transmitted positions within the Chinese EEZ. This may point to vessel 
changing to official MMSI to enter the Chinese EEZ.  


Now lets try plot out some of this effort data to share with JAC partners 

```{r warning=FALSE}
# lets aggregate the data to lat/lon bins to use fishwatchr example mapping code 
voi_ais$lat_bin <- round(voi_ais$lat,1)
voi_ais$lon_bin <- round(voi_ais$lon,1)

presence <- voi_ais %>%
  group_by(lat_bin, lon_bin) %>%
    summarise(
      positions = n(), 
    )

  presence %>%
  recenter_raster(raster_df = .,
                  res = 1,
                  x_lab = 'lon_bin',
                  y_lab = 'lat_bin',
                  fill_lab = 'positions',
                  center = -80) %>%
  ggplot() +
  geom_raster(aes(x = lon_bin,
                  y = lat_bin,
                  fill = positions)) +
  geom_gfw_land(theme = 'light', center = -80) +
  geom_gfw_eez(theme = 'light', center = -80, alpha = 0.05) +
  scale_fill_gradientn(colors = gfw_palette('map_effort_dark'),
                       limits = c(0,10000),
                       oob = scales::squish,
                       na.value = NA) +
  labs(fill = 'AIS Positions') +
  theme_gfw_map(theme = 'light')

ggsave(here::here("outputs", "figures", "voi_ais_map.png"))

```


Save the data in the repo to save pulling from big query each time 
```{r}
voi_ais %>% write_csv(here::here("data", ".", "voi_ais.csv"))
voi_ais %>% saveRDS(here::here("data", ".", "voi_ais.rds"))
f_eff <- read_rds(here::here("data", ".", "voi_ais.rds"))
```

Now test out a regional map with machine learning derived fishing events 

```{r, echo=F}
mmsi_412331284 <- c("
SELECT 
  * 
FROM `world-fishing-827.pipe_production_v20201001.messages_segmented_20221111`
WHERE 
  ssvid = '412331284' AND 
  DATE(timestamp) BETWEEN '2020-11-01' AND DATE(CURRENT_DATE())")


mmsi_412331284_ais <- fishwatchr::gfw_query(query = mmsi_412331284,
                                 run_query = TRUE, 
                                 con = con)$data
head(mmsi_412331284_ais)

mmsi_412331284_ais$regions[1]$eez


out_dat <- dplyr::select(mmsi_412331284_ais, source, type, ssvid, timestamp, 
                        lon, lat, speed, course, heading, shipname, callsign, 
                        destination, imo, shiptype, receiver_type, receiver, 
                        length, width, status)
write.csv(out_dat, 'LRYY715 412331284.csv', row.names = F)

mmsi_150400453 <- c("
SELECT 
  * 
FROM `world-fishing-827.pipe_production_v20201001.messages_scored_20221111`
WHERE 
  ssvid = '150400453' AND 
  DATE(timestamp) BETWEEN '2020-11-01' AND DATE(CURRENT_DATE())")


mmsi_150400453_ais  <- fishwatchr::gfw_query(query = mmsi_150400453,
                                 run_query = TRUE, 
                                 con = con)$data
head(mmsi_150400453_ais)


out_dat2 <- dplyr::select(mmsi_150400453_ais, source, type, ssvid, timestamp, 
                        lon, lat, speed, course, heading, shipname, callsign, 
                        destination, imo, shiptype, receiver_type, receiver, 
                        length, width, status)
write.csv(out_dat2, 'LRYY715 150400453.csv', row.names = F)

```

