---
title: "Bahia Posesion - Chile"
author: "Max Schofield"
date: "2022-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

foo <- function(x){
  for( i in x ){
    #  require returns TRUE invisibly if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      #  Load package after installing
      require( i , character.only = TRUE )
    }
  }
}

#  Then try/install packages...
foo( c("tidyverse" , "bigrquery" ,"devtools", "DBI","glue", "lubridate", "here", "sf", "extrafont", "patchwork", "terra") )

# get fishwatch r independently 
if (!require("fishwatchr")) {devtools::install_github("GlobalFishingWatch/fishwatchr")}
library(fishwatchr)

```

```{r con}
con <- DBI::dbConnect(drv = bigrquery::bigquery(), 
                      project = "world-fishing-827", 
                      use_legacy_sql = FALSE)
```

## Brief 

Joint Analytical Cell has been conducting intelligence reports (JAC X,  JAC Y) on 
Chinese flagged squid vessels that are operating in the southwest Atlantic and 
southeast Pacific fishing grounds. The vessels were the subject of JAC intelligence 
because they consistently change identity on AIS in an apparent effort to hide their 
activities. These vessels have been transiting between the Atlantic and Pacific 
fishing grounds through the Straight of Magellan, Chile where on route they often 
stop in Bahia Posesión, Chile.

This analysis investigates the vessel activity in Bahia Posesion to try and ascertain
the who, where based on GFW AIS data. 

```{r}
# define bounding box/area of interest
bbox <- data.frame(x_min = -69.40,
                   x_max = -69.13,
                   y_min = -52.31,
                   y_max = -52.20)
# what projection should be used for mapping
best_proj <- fishwatchr::gfw_projections("Equal Earth")$proj_string
# best_proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```


```{r}
bp_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_vessel_info.sql"))

bp <- fishwatchr::gfw_query(query = bp_q,
                                 run_query = TRUE,
                                 con = con)$data

# write.csv(distinct(dplyr::select(bp, ssvid, best_flag, best_vessel_class, vessel_name)), 'Bahia_Posession_2022.csv')
write.csv(distinct(dplyr::select(bp, ssvid, best_flag, best_vessel_class, vessel_name)), 'Bahia_Posession_2021.csv')

```

Vessel activity in the Bahia Posesion peaked had distinct peaks May 2021 and June 2022. 
There is very little activity on AIS in Bahia Posesion outside of these periods. 


```{r}
# get rid of duplicate values based on name / auths 
bp_ves_num <- bp[!duplicated(bp[,1:5]), c(1:5)]

# make month a factor to allow values for missing levels for plotting 
bp_ves_num$month <- factor(bp_ves_num$month, levels = c(1:12))

# create a complete df for showing vessel movements through time 
plot_df <- bp_ves_num %>% group_by(year, month)  %>% summarise(
  vessels = n_distinct(ssvid)
) %>% complete(month)

# make NAs 0s 
plot_df$vessels[is.na(plot_df$vessels)] <- 0

# make 2022 12 an NA as its in the future  
plot_df$vessels[plot_df$year==2022 & plot_df$month == 12] <- NA

# plot showing vessel numbers through time - gfw theme 
p1 <- ggplot(plot_df)+
  geom_point(aes(x=month, y=vessels, colour=factor(year)))+
  geom_line(aes(x=as.numeric(as.character(month)), y=vessels, colour=factor(year))) +
  scale_x_discrete(breaks=1:12,labels=c('Jan', 'Feb', 'Mar', 'Apr', 'May', 
                      'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec'))+
  scale_colour_manual(values=gfw_palette('primary')) +
  labs(x='Month', y='Vessels', colour='Year')+
  theme_gfw() +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold"))

# run this direct in console 
# setwd('./outputs/figures')
ggsave('Bahia_posesion_vessels_2021_2022.png', plot=p1, device='png')

```
Look at the length of stay in the area 

```{r}
dates_bp_q <- c("
SELECT 
  DISTINCT ssvid,
  DATE(timestamp) AS date, 
  EXTRACT(year FROM timestamp) AS year,
FROM `world-fishing-827.scratch_max.bahia_posesion_raw_ais`
")


dates_bp <- fishwatchr::gfw_query(query = dates_bp_q,
                                 run_query = TRUE, 
                                 con = con)$data

plot_df2 <- dates_bp %>% group_by(ssvid, year) %>% summarise(days = n_distinct(date))

p2 <- ggplot(plot_df2)+
  geom_bar(aes(x=days, fill=factor(year)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('primary')) +
  labs(x='Days spent in Bahia Posesion', y='Frequency', fill='Year')+
  theme_gfw()+
  theme(axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"))

# run this direct in console 
ggsave('Bahia_posesion_length_of_stay_2021_2022.png',plot=p2, device='png')

```

Look at what gears are getting used. 

```{r}
# plot of the gears 

# set NA to unknown
bp_ves_num$best_vessel_class[is.na(bp_ves_num$best_vessel_class)] <- 'unknown'

p2 <- ggplot(bp_ves_num)+
  geom_bar(aes(x=best_vessel_class, fill=factor(year)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('primary')) +
  labs(x='Vessel Type', y='Vessels', fill='Year')+
  theme_gfw()+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"), 
        legend.title=element_text(size=12), legend.text=element_text(size=11)) 

# run this direct in console 
ggsave('Bahia_posesion_vessel_type_2021_2022.png', device='png')

```
Look at what flags the vessels are flying. 

```{r}
# plot of the gears 

# set NA to unknown
bp_ves_num$best_flag[is.na(bp_ves_num$best_flag)] <- 'unknown'

# just keep ssvid, flag and vessel type 
plot_df3 <- bp_ves_num[!duplicated(bp_ves_num[,c(1,4,5)]),c(1,4,5)]

p3 <- ggplot(plot_df3)+
  geom_bar(aes(x=best_vessel_class, fill=factor(best_flag)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('chart')) +
  labs(x='Vessel Type', y='Vessels', fill='Flag State')+
  theme_gfw()+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"), 
        legend.title=element_text(size=12), legend.text=element_text(size=11)) 

# run this direct in console 
ggsave('Bahia_posesion_vessel_type_flag_2021_2022.png', device='png')
```

Look at the vessels authorisations

```{r}
auth_bp_q <- readr::read_file(file = here::here("queries", ".", "authorisations_bp.sql"))

auth_bp <- fishwatchr::gfw_query(query = auth_bp_q,
                                 run_query = TRUE,
                                 con = con)$data

ggplot(auth_bp) +
  geom_bar(aes(x=authorisation))
auth_bp %>% 
  group_by(authorisation) %>% 
  summarise(count=n()) %>%
  plot_donut_chart(group_var = authorisation,
                   value_var = count,
                   donut_title = 'Authorisation level of fleet') +
  labs(fill = 'Known Authorisations Sources')
   

```

Look for any high risk vessels

```{r}


```
