---
title: "Bahia Posesion - Chile"
author: "Max Schofield"
date: "2022-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

foo <- function(x){
  for( i in x ){
    #  require returns TRUE invisibly if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      #  Load package after installing
      require( i , character.only = TRUE )
    }
  }
}

#  Then try/install packages...
foo( c("tidyverse" , "bigrquery" ,"devtools", "DBI","glue", "lubridate", "here", "sf", "extrafont", "patchwork", "terra") )

# get fishwatch r independently 
if (!require("fishwatchr")) {devtools::install_github("GlobalFishingWatch/fishwatchr")}
library(fishwatchr)

```

```{r con}
con <- DBI::dbConnect(drv = bigrquery::bigquery(), 
                      project = "world-fishing-827", 
                      use_legacy_sql = FALSE)
```

## Brief 

Joint Analytical Cell has been conducting intelligence reports (JAC X,  JAC Y) on 
Chinese flagged squid vessels that are operating in the southwest Atlantic and 
southeast Pacific fishing grounds. The vessels were the subject of JAC intelligence 
because they consistently change identity on AIS in an apparent effort to hide their 
activities. These vessels have been transiting between the Atlantic and Pacific 
fishing grounds through the Straight of Magellan, Chile where on route they often 
stop in Bahia Posesión, Chile.

This analysis investigates the vessel activity in Bahia Posesion to try and ascertain
the who, where based on GFW AIS data. 

```{r}
# define bounding box/area of interest
bbox <- data.frame(x_min = -69.40,
                   x_max = -69.13,
                   y_min = -52.31,
                   y_max = -52.20)
# what projection should be used for mapping
best_proj <- fishwatchr::gfw_projections("Equal Earth")$proj_string
# best_proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```


```{r}
bp_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_vessel_info.sql"))

bp <- fishwatchr::gfw_query(query = bp_q,
                                 run_query = TRUE,
                                 con = con)$data

# write.csv(distinct(dplyr::select(bp, ssvid, best_flag, best_vessel_class, vessel_name)), 'Bahia_Posession_2022.csv')
write.csv(distinct(dplyr::select(bp, ssvid, best_flag, best_vessel_class, vessel_name)), 'Bahia_Posession_2021.csv')

```

Vessel activity in the Bahia Posesion peaked had distinct peaks May 2021 and June 2022. 
There is very little activity on AIS in Bahia Posesion outside of these periods. 


```{r}
# get rid of duplicate values based on name / auths 
bp_ves_num <- bp[!duplicated(bp[,1:5]), c(1:5)]

# make month a factor to allow values for missing levels for plotting 
bp_ves_num$month <- factor(bp_ves_num$month, levels = c(1:12))

# create a complete df for showing vessel movements through time 
plot_df <- bp_ves_num %>% group_by(year, month)  %>% summarise(
  vessels = n_distinct(ssvid)
) %>% complete(month)

# make NAs 0s 
plot_df$vessels[is.na(plot_df$vessels)] <- 0

# make 2022 12 an NA as its in the future  
plot_df$vessels[plot_df$year==2022 & plot_df$month == 12] <- NA

# plot showing vessel numbers through time - gfw theme 
p1 <- ggplot(plot_df)+
  geom_point(aes(x=month, y=vessels, colour=factor(year)))+
  geom_line(aes(x=as.numeric(as.character(month)), y=vessels, colour=factor(year))) +
  scale_x_discrete(breaks=1:12,labels=c('Jan', 'Feb', 'Mar', 'Apr', 'May', 
                      'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec'))+
  scale_colour_manual(values=gfw_palette('primary')) +
  labs(x='Month', y='Vessels', colour='Year')+
  theme_gfw() +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold"))

# run this direct in console 
# setwd('./outputs/figures')
ggsave('Bahia_posesion_vessels_2021_2022.png', plot=p1, device='png')

```
Look at the length of stay in the area 

```{r}
dates_bp_q <- c("
SELECT 
  DISTINCT ssvid,
  DATE(timestamp) AS date, 
  EXTRACT(year FROM timestamp) AS year,
FROM `world-fishing-827.scratch_max.bahia_posesion_raw_ais`
")


dates_bp <- fishwatchr::gfw_query(query = dates_bp_q,
                                 run_query = TRUE, 
                                 con = con)$data

plot_df2 <- dates_bp %>% group_by(ssvid, year) %>% summarise(days = n_distinct(date))

p2 <- ggplot(plot_df2)+
  geom_bar(aes(x=days, fill=factor(year)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('primary')) +
  labs(x='Days spent in Bahia Posesion', y='Frequency', fill='Year')+
  theme_gfw()+
  theme(axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"))

# run this direct in console 
ggsave('Bahia_posesion_length_of_stay_2021_2022.png',plot=p2, device='png')

```

Look at what gears are getting used. 

```{r}
# plot of the gears 

# set NA to unknown
bp_ves_num$best_vessel_class[is.na(bp_ves_num$best_vessel_class)] <- 'unknown'

p2 <- ggplot(bp_ves_num)+
  geom_bar(aes(x=best_vessel_class, fill=factor(year)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('primary')) +
  labs(x='Vessel Type', y='Vessels', fill='Year')+
  theme_gfw()+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"), 
        legend.title=element_text(size=12), legend.text=element_text(size=11)) 

# run this direct in console 
ggsave('Bahia_posesion_vessel_type_2021_2022.png', device='png')

```
Look at what flags the vessels are flying. 

```{r}
# plot of the gears 

# set NA to unknown
bp_ves_num$best_flag[is.na(bp_ves_num$best_flag)] <- 'unknown'

# just keep ssvid, flag and vessel type 
plot_df3 <- bp_ves_num[!duplicated(bp_ves_num[,c(1,4,5)]),c(1,4,5)]

p3 <- ggplot(plot_df3)+
  geom_bar(aes(x=best_vessel_class, fill=factor(best_flag)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('chart')) +
  labs(x='Vessel Type', y='Vessels', fill='Flag State')+
  theme_gfw()+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"), 
        legend.title=element_text(size=12), legend.text=element_text(size=11)) 

# run this direct in console 
ggsave('Bahia_posesion_vessel_type_flag_2021_2022.png', device='png')
```

Look at the vessels authorisations

```{r}
auth_bp_q <- readr::read_file(file = here::here("queries", ".", "authorisations_bp.sql"))

auth_bp <- fishwatchr::gfw_query(query = auth_bp_q,
                                 run_query = TRUE,
                                 con = con)$data


# only look for auths for fishing vessels
fishing <- filter(auth_bp, best_vessel_class %in% c('fishing', 'squid_jigger'))

n_distinct(fishing$ssvid)

sprfmo <- filter(fishing, registry %in% c('CHINASPRFMO', 'SPRFMO', 'SPRFMO2'))
n_distinct(sprfmo$ssvid)

fishing[!fishing$ssvid %in% sprfmo$ssvid,]

filter(fishing, ssvid == '412549299')


# ggplot(auth_bp) +
#   geom_bar(aes(x=authorisation))
# auth_bp %>% 
#   group_by(authorisation) %>% 
#   summarise(count=n()) %>%
#   plot_donut_chart(group_var = authorisation,
#                    value_var = count,
#                    donut_title = 'Authorisation level of fleet') +
#   labs(fill = 'Known Authorisations Sources')
#    

```

Show vessel movements to Bahia Posesion and after leaving Bahia Posesion 
```{r}
move_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_voi_squid_location_2021_2022.sql"))

move <- fishwatchr::gfw_query(query = move_q,
                                 run_query = TRUE,
                                 con = con)$data

# need a positions dataset with EEZ
# something like this should do the job 
eez_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_vessels_eez_activity_2021_2022.sql"))

eez_act <- fishwatchr::gfw_query(query = eez_q,
                                 run_query = TRUE,
                                 con = con)$data

# group the EEZs into informative larger groups 
# these might need to go into EEZ level still to show movement to and from Bahia Posesion 
table(eez_act$eez)

eez_act$eez_group <- eez_act$eez
eez_act$eez_group[eez_act$eez %in% c('IND','SOM','YEM','LKA','MDV','OMN','DJI',
                                     'IRN','PAK','ERI','ERI-DJI','SAU','ARE')] <- 'nw_indian_eezs'
eez_act$eez_group[eez_act$eez %in% c('MUS', 'MDG', 'ZAF', 'MYS')] <- 'sw_indian_eezs'
eez_act$eez_group[eez_act$eez %in% c('CHL','PER','ECU')] <- 'se_pacific_eezs'
eez_act$eez_group[eez_act$eez %in% c('ARG','FLK','URY','BRA')] <- 'sw_atlantic_eezs'
eez_act$eez_group[eez_act$eez %in% c('JPN','KOR','PHL','MNP','TWN','TWN-JPN-CHN',
                                     'JPN-RUS','RUS','PRK','CHN','JPN-KOR')] <- 'nw_pacific_eezs'
eez_act$eez_group[!eez_act$eez_group %in% c('nw_indian_eezs', 'se_pacific_eezs', 
                                            'sw_atlantic_eezs', 'nw_pacific_eezs', 
                                            'sw_indian_eezs')] <- 'Other'

# use Global Squid paper pipes to make plotting dfs 

# use aoi tagged data and identify when vessel moves between aois 'change aoi' 
base_data <- move %>%
  group_by(ssvid) %>%
  arrange(date) %>%
  mutate(change_aoi = c(0, cumsum(aoi[-1L] != aoi[-length(aoi)] |
    as.numeric(difftime(date[-1L],
      date[-length(date)],
      units = "day"
    )) > 14))) %>%
  ungroup() %>%
  dplyr::select(ssvid, aoi, date, change_aoi, everything()) %>%
  ungroup()

# find the start and end points of the change in aoi for segmenting 
time_range_data <- move %>%
  arrange(date) %>%
  group_by(ssvid) %>%
  mutate(change_aoi = c(0, cumsum(aoi[-1L] != aoi[-length(aoi)] |
    as.numeric(difftime(date[-1L],
      date[-length(date)],
      units = "day"
    )) > 14))) %>%
  ungroup() %>%
  dplyr::select(ssvid, aoi, date, change_aoi, everything()) %>%
  group_by(ssvid, change_aoi) %>%
  summarise(
    first_date = min(date),
    last_date = max(date)
  ) %>%
  ungroup()

# join two datasets together 
merged_data <- base_data %>%
  inner_join(time_range_data %>%
    dplyr::select(ssvid, first_date, last_date),
  by = c("ssvid" = "ssvid", "date" = "first_date")
  ) %>%
  dplyr::select(ssvid, aoi, date, last_date, everything()) %>%
  rename(first_date = date) %>%
  # rbind(aoi_padding) %>%
  mutate(aoi3 = factor(aoi, levels = c(
    "nw_pacific","se_pacific", "sw_atlantic", "nw_indian",
    "nw_pacific_eezs","se_pacific_eezs","sw_atlantic_eezs","nw_indian_eezs"
  ))) %>%
  filter(!is.na(ssvid))

# need to establish the purpose of merged 3. Could make chunks comparable to merge 
# using the activity data within EEZs and join them together 

# don't need to find most common aoi as all vessels in same group. 
# need to pull together 
eez_activity <- eez_act %>%
  arrange(date) %>%
  mutate(change_aoi = c(0, cumsum(eez_group[-1L] != eez_group[-length(eez_group)]))) %>%
  group_by(ssvid, eez_group, change_aoi) %>%
  summarize(
    min_date = min(date),
    max_date = max(date),
    duration = as.numeric(difftime(max_date, min_date, tz = "UTC", "secs")),
    duration = ifelse(min_date == max_date, 86400, duration)
  ) %>%
  ungroup() %>%
  arrange(min_date) %>%
  group_by(ssvid, eez_group) %>%
  summarize(total_days = sum(duration)) %>%
  ungroup() %>%
  group_by(ssvid) %>%
  filter(total_days == max(total_days)) %>%
  select(ssvid, most_common_aoi = eez_group) #%>%
  # inner_join(merged_data,
  #   by = c("ssvid" = "ssvid")
  # )

merged_data3 <- merged_data2 %>%
  mutate(
    ssvid_order = paste0(most_common_aoi, ssvid)
  ) %>%
  mutate(ssvid = factor(ssvid, levels = sort(voyages_ssvids)))

segment_data <- merged_data3 %>%
  mutate(
    ssvid_order = paste0(most_common_aoi, ssvid)
  ) %>%
  group_by(ssvid, flag) %>%
  summarize(
    start = min(first_date, na.rm = TRUE),
    end = max(last_date, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(ssvid = factor(ssvid, levels = sort(voyages_ssvids)))

p <- ggplot() +
  geom_segment(
    data = segment_data,
    aes(
      x = start,
      xend = end,
      y = ssvid,
      yend = ssvid
    ),
    color = "grey90"
  ) +
  geom_segment(
    data = merged_data3,
    aes(
      x = first_date,
      xend = last_date,
      y = ssvid,
      yend = ssvid,
      # group = ssvid_order,
      color = aoi3
    ),
    # color = 'grey20',
    size = {{ line_width }},
    alpha = 1
  ) +
  scale_color_manual("Fishing Region",
    breaks = c(
      "nw_pacific","se_pacific", "sw_atlantic", "nw_indian",
      "nw_pacific_eezs","se_pacific_eezs","sw_atlantic_eezs","nw_indian_eezs"
    ),
    values = c("#5D69B1", "#58E8C6", "#99C945", "#CC3A8E",
               "#465192", "#148f72","#587622","#822259"),
    labels = c(
      "NW Pacific Ocean High Seas", "SE Pacific Ocean High Seas",
      "SW Atlantic Ocean High Seas", "NW Indian Ocean High Seas",
      "NW Pacific Ocean EEZs", "SE Pacific Ocean EEZs",
      "SW Atlantic Ocean EEZs", "NW Indian Ocean EEZs"
    )
  ) +
  scale_x_date("",
    date_minor_breaks = "months",
    date_breaks = "year",
    date_labels = "%Y",
    limits = c(as.Date("2017-01-01"), as.Date("2020-12-31")),
    expand = c(0, 0)
  ) +
  theme_gfw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.y = element_text(size = 1),
    #axis.text.y = element_blank(),
    panel.border = element_rect(color = "grey85", fill = NA),
    panel.background = element_rect(fill = "#ffffff"),
    plot.background = element_rect(fill = "#ffffff"),
    strip.text.y.left = element_text(angle = 0),
    panel.grid.minor.x = element_line(color = "grey85", linetype = "dotted"),
    axis.title = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(family = "Roboto-Regular")
  ) 
  p

```


```{r}
bp_enc_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_encounters.sql"))

# bp_enc <- fishwatchr::gfw_query(query = glue::glue(bp_enc_q,
#                                               date_start = '2021-01-01',
#                                               date_end = '2022-11-30'),
bp_enc <- fishwatchr::gfw_query(query = bp_enc_q,
                                 run_query = TRUE,
                                 con = con)$data

```
```{r}

bp_q <- c("
SELECT 
  *
FROM `world-fishing-827.scratch_max.bahia_posesion_raw_ais`
")


bp_q <- fishwatchr::gfw_query(query = bp_q,
                                 run_query = TRUE, 
                                 con = con)$data

write.csv(bp_q,'data/BP_raw_ais.csv')

```
