---
title: "Bahia Posesion - Chile"
author: "Max Schofield"
date: "2022-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

foo <- function(x){
  for( i in x ){
    #  require returns TRUE invisibly if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      #  Load package after installing
      require( i , character.only = TRUE )
    }
  }
}

#  Then try/install packages...
foo( c("tidyverse" , "bigrquery" ,"devtools", "DBI","glue", "lubridate", "here", "sf", "extrafont", "patchwork", "terra") )

# get fishwatch r independently 
if (!require("fishwatchr")) {devtools::install_github("GlobalFishingWatch/fishwatchr")}
library(fishwatchr)

```

```{r con}
con <- DBI::dbConnect(drv = bigrquery::bigquery(), 
                      project = "world-fishing-827", 
                      use_legacy_sql = FALSE)
```

## Brief 

Joint Analytical Cell has been conducting intelligence reports (JAC X,  JAC Y) on 
Chinese flagged squid vessels that are operating in the southwest Atlantic and 
southeast Pacific fishing grounds. The vessels were the subject of JAC intelligence 
because they consistently change identity on AIS in an apparent effort to hide their 
activities. These vessels have been transiting between the Atlantic and Pacific 
fishing grounds through the Straight of Magellan, Chile where on route they often 
stop in Bahia Posesión, Chile.

This analysis investigates the vessel activity in Bahia Posesion to try and ascertain
the who, where based on GFW AIS data. 

```{r}
# define bounding box/area of interest
bbox <- data.frame(x_min = -69.40,
                   x_max = -69.13,
                   y_min = -52.31,
                   y_max = -52.20)
# what projection should be used for mapping
best_proj <- fishwatchr::gfw_projections("Equal Earth")$proj_string
# best_proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```


```{r}
bp_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_vessel_info.sql"))

bp <- fishwatchr::gfw_query(query = bp_q,
                                 run_query = TRUE,
                                 con = con)$data

# write.csv(distinct(dplyr::select(bp, ssvid, best_flag, best_vessel_class, vessel_name)), 'Bahia_Posession_2022.csv')
write.csv(distinct(dplyr::select(bp, ssvid, best_flag, best_vessel_class, vessel_name)), 'Bahia_Posession_2021.csv')

```

Vessel activity in the Bahia Posesion peaked had distinct peaks May 2021 and June 2022. 
There is very little activity on AIS in Bahia Posesion outside of these periods. 


```{r}
# get rid of duplicate values based on name / auths 
bp_ves_num <- bp[!duplicated(bp[,1:5]), c(1:5)]

# make month a factor to allow values for missing levels for plotting 
bp_ves_num$month <- factor(bp_ves_num$month, levels = c(1:12))

# create a complete df for showing vessel movements through time 
plot_df <- bp_ves_num %>% group_by(year, month)  %>% summarise(
  vessels = n_distinct(ssvid)
) %>% complete(month)

# make NAs 0s 
plot_df$vessels[is.na(plot_df$vessels)] <- 0

# make 2022 12 an NA as its in the future  
plot_df$vessels[plot_df$year==2022 & plot_df$month == 12] <- NA

# plot showing vessel numbers through time - gfw theme 
p1 <- ggplot(plot_df)+
  geom_point(aes(x=month, y=vessels, colour=factor(year)))+
  geom_line(aes(x=as.numeric(as.character(month)), y=vessels, colour=factor(year))) +
  scale_x_discrete(breaks=1:12,labels=c('Jan', 'Feb', 'Mar', 'Apr', 'May', 
                      'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec'))+
  scale_colour_manual(values=gfw_palette('primary')) +
  labs(x='Month', y='Number of Vessels', colour='Year')+
  theme_gfw() +
  theme(axis.text=element_text(size=11), axis.title=element_text(size=13,face="bold"),
        legend.title=element_text(size=12), legend.text=element_text(size=11))

getwd()
# run this direct in console 
# setwd('./outputs/figures')
ggsave('../outputs/figures/Bahia_posesion_vessels_2021_2022.png', plot=p1, device='png')

```
Look at the length of stay in the area 

```{r}
dates_bp_q <- c("
SELECT 
  DISTINCT ssvid,
  DATE(timestamp) AS date, 
  EXTRACT(year FROM timestamp) AS year,
FROM `world-fishing-827.scratch_max.bahia_posesion_raw_ais`
")


dates_bp <- fishwatchr::gfw_query(query = dates_bp_q,
                                 run_query = TRUE, 
                                 con = con)$data

plot_df2 <- dates_bp %>% group_by(ssvid, year) %>% summarise(days = n_distinct(date))

p2 <- ggplot(plot_df2)+
  geom_bar(aes(x=days, fill=factor(year)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('primary')) +
  labs(x='Days spent in Bahia Posesion', y='Number of vessels', fill='Year')+
  theme_gfw()+
  theme(axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"),
        legend.title=element_text(size=12), legend.text=element_text(size=11))

# run this direct in console 
ggsave('../outputs/figures/Bahia_posesion_length_of_stay_2021_2022.png',plot=p2, device='png')

```

Look at what gears are getting used. 

```{r}
# plot of the gears 

# set NA to unknown
bp_ves_num$best_vessel_class[is.na(bp_ves_num$best_vessel_class)] <- 'unknown'

# look at class options 
table(bp_ves_num$best_vessel_class)


# corrections of data where spurious MMSIs are getting misclassified 
bp_ves_num$best_vessel_class[bp_ves_num$ssvid=='210000000'] <- 'unknown'
bp_ves_num$best_flag[bp_ves_num$ssvid=='210000000'] <- 'unknown'
bp_ves_num$best_vessel_class[bp_ves_num$ssvid=='499999999'] <- 'unknown'
bp_ves_num$best_flag[bp_ves_num$ssvid=='499999999'] <- 'unknown'


# make a plot vessel class 
bp_ves_num$plot_vessel_class <- bp_ves_num$best_vessel_class
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='cargo_or_reefer'] <- 'Cargo or Reefer'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='fishing'] <- 'Fishing'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='gear'] <- 'Gear'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='non_fishing'] <- 'Non Fishing'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='passenger'] <- 'Passenger'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='pole_and_line'] <- 'Pole and Line'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='squid_jigger'] <- 'Squid Jigger'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='supply_vessel'] <- 'Supply Vessel'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='tanker'] <- 'Tanker'
bp_ves_num$plot_vessel_class[bp_ves_num$plot_vessel_class=='unknown'] <- 'Unknown'


p2 <- ggplot(bp_ves_num)+
  geom_bar(aes(x=plot_vessel_class, fill=factor(year)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('primary')) +
  labs(x='Vessel Type', y='Number of Vessels', fill='Year')+
  theme_gfw()+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"), 
        legend.title=element_text(size=12), legend.text=element_text(size=11)) 

# run this direct in console 
ggsave('../outputs/figures/Bahia_posesion_vessel_type_2021_2022.png', device='png')

```
Look at what flags the vessels are flying. 

```{r}
# plot of the gears 

# set NA to unknown
bp_ves_num$best_flag[is.na(bp_ves_num$best_flag)] <- 'Unknown'
bp_ves_num$full_flag <- bp_ves_num$best_flag
table(bp_ves_num$full_flag)

bp_ves_num$full_flag[bp_ves_num$full_flag=='CHL'] <- 'Chile'
bp_ves_num$full_flag[bp_ves_num$full_flag=='CHN'] <- 'China'
bp_ves_num$full_flag[bp_ves_num$full_flag=='IMN'] <- 'Isle of Man'
bp_ves_num$full_flag[bp_ves_num$full_flag=='LBR'] <- 'Liberia'
bp_ves_num$full_flag[bp_ves_num$full_flag=='MHL'] <- 'Marshall Islands'
bp_ves_num$full_flag[bp_ves_num$full_flag=='unknown'] <- 'Unknown'

# just keep ssvid, flag and vessel type 
plot_df3 <- bp_ves_num[!duplicated(bp_ves_num[,c(1,6,7)]),c(1,6,7)]

p3 <- ggplot(plot_df3)+
  geom_bar(aes(x=plot_vessel_class, fill=factor(full_flag)), 
           position = position_dodge2(preserve = 'single')) +
  scale_fill_manual(values=gfw_palette('chart')) +
  labs(x='Vessel Type', y='Vessels', fill='Flag State')+
  theme_gfw()+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), 
        axis.text=element_text(size=11), axis.title=element_text(size=12,face="bold"), 
        legend.title=element_text(size=12), legend.text=element_text(size=11)) 

# run this direct in console 
ggsave('../outputs/figures/Bahia_posesion_vessel_type_flag_2021_2022.png', device='png')
```

Look at the vessels authorisations

```{r}
auth_bp_q <- readr::read_file(file = here::here("queries", ".", "authorisations_bp.sql"))

auth_bp <- fishwatchr::gfw_query(query = auth_bp_q,
                                 run_query = TRUE,
                                 con = con)$data


auth_bp$sprfmo <- ifelse(auth_bp$registry %in% c('CHINASPRFMO', 'SPRFMO', 'SPRFMO2'), TRUE, FALSE)

working_df <- auth_bp %>% filter(!best_vessel_class %in% c('carfo_reefer', 'non_fishing', 'passenger', 'supply_vessel', 'tanker')) %>%
                        group_by(ssvid) %>% 
                            summarise(
                                flag = first(best_flag), 
                                vessel_class = first(best_vessel_class), 
                                sprfmo = any(sprfmo))

# look at vessels not in sprfmo
working_df %>% filter(!sprfmo)

# drop out non fishing vessels 
working_df <- working_df %>% filter(!ssvid %in% c('949403626', '725003591', '725000317', '636020482'))

# table(working_df$vessel_class)
# 
# # working_df <- bp[,c(1,5,6)] %>% filter(!duplicated(ssvid))
# # n_distinct(working_df$ssvid)
# 
# table(working_df$vessel_name)
# 
# filter(working_df, is.na(best_vessel_class)) 
# 
# # only look for auths for fishing vessels
# fishing <- filter(auth_bp, best_vessel_class %in% c('fishing', 'squid_jigger'))
# 
# n_distinct(fishing$ssvid)
# 
# sprfmo <- filter(fishing, registry %in% c('CHINASPRFMO', 'SPRFMO', 'SPRFMO2'))
# n_distinct(sprfmo$ssvid)
# 
# fishing[!fishing$ssvid %in% sprfmo$ssvid,]
# 
# filter(fishing, ssvid == '412549299')
# 
# 
# # ggplot(auth_bp) +
# #   geom_bar(aes(x=authorisation))
# # auth_bp %>% 
# #   group_by(authorisation) %>% 
# #   summarise(count=n()) %>%
# #   plot_donut_chart(group_var = authorisation,
# #                    value_var = count,
# #                    donut_title = 'Authorisation level of fleet') +
# #   labs(fill = 'Known Authorisations Sources')
# #    

```

Show vessel movements to Bahia Posesion and after leaving Bahia Posesion 
```{r}

# import in the movement data from BQ
move_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_voi_squid_location_2021_2022.sql"))

move <- fishwatchr::gfw_query(query = move_q,
                                 run_query = TRUE,
                                 con = con)$data

# import the activity data in EEZs
eez_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_vessels_eez_activity_2021_2022.sql"))

eez_act <- fishwatchr::gfw_query(query = eez_q,
                                 run_query = TRUE,
                                 con = con)$data

# group the EEZs into informative larger groups 
# these might need to go into EEZ level still to show movement to and from Bahia Posesion 
table(eez_act$eez)

eez_act$eez_group <- eez_act$eez
# reclassifying NW Indian Ocean to Other EEZ 
eez_act$eez_group[eez_act$eez %in% c('IND','SOM','YEM','LKA','MDV','OMN','DJI',
                                     'IRN','PAK','ERI','ERI-DJI','SAU','ARE')] <- 'other_eez'
eez_act$eez_group[eez_act$eez %in% c('MUS', 'MDG', 'ZAF', 'MYS')] <- 'sw_indian_eezs'
eez_act$eez_group[eez_act$eez %in% c('CHL','PER','ECU')] <- 'se_pacific_eezs'
eez_act$eez_group[eez_act$eez %in% c('ARG','FLK','URY','BRA')] <- 'sw_atlantic_eezs'
eez_act$eez_group[eez_act$eez %in% c('JPN','KOR','PHL','MNP','TWN','TWN-JPN-CHN',
                                     'JPN-RUS','RUS','PRK','CHN','JPN-KOR')] <- 'nw_pacific_eezs'
eez_act$eez_group[!eez_act$eez_group %in% c('nw_indian_eezs', 'se_pacific_eezs', 
                                            'sw_atlantic_eezs', 'nw_pacific_eezs', 
                                            'sw_indian_eezs')] <- 'other_eez'

# use Global Squid paper pipes to make plotting dfs 
# use aoi tagged data and identify when vessel moves between aois 'change aoi' 
base_data <- move %>%
  group_by(ssvid) %>%
  arrange(date) %>%
  mutate(change_aoi = c(0, cumsum(aoi[-1L] != aoi[-length(aoi)] |
    as.numeric(difftime(date[-1L],
      date[-length(date)],
      units = "day"
    )) > 14))) %>%
  ungroup() %>%
  dplyr::select(ssvid, aoi, date, change_aoi, everything()) %>%
  ungroup()

# find the start and end points of the change in aoi for segmenting 
time_range_data <- move %>%
  arrange(date) %>%
  group_by(ssvid) %>%
  mutate(change_aoi = c(0, cumsum(aoi[-1L] != aoi[-length(aoi)] |
    as.numeric(difftime(date[-1L],
      date[-length(date)],
      units = "day"
    )) > 14))) %>%
  ungroup() %>%
  dplyr::select(ssvid, aoi, date, change_aoi, everything()) %>%
  group_by(ssvid, change_aoi) %>%
  summarise(
    first_date = min(date),
    last_date = max(date)
  ) %>%
  ungroup()

# join two datasets together 
merged_data <- base_data %>%
  inner_join(time_range_data %>%
    dplyr::select(ssvid, first_date, last_date),
  by = c("ssvid" = "ssvid", "date" = "first_date")
  ) %>%
  dplyr::select(ssvid, aoi, date, last_date, everything()) %>%
  rename(first_date = date) %>%
  # rbind(aoi_padding) %>%
  mutate(aoi3 = factor(aoi, levels = c(
    "nw_pacific","se_pacific", "sw_atlantic", "nw_indian",
    "other_eez","sw_indian_eezs","bahia_posesion",
    "nw_pacific_eezs","se_pacific_eezs","sw_atlantic_eezs"
  ))) %>%
  filter(!is.na(ssvid))

#------------------------------
# repeat for EEZ activity 
# use aoi tagged data and identify when vessel moves between aois 'change aoi' 
base_data2 <- eez_act %>%
  group_by(ssvid) %>%
  arrange(date) %>%
  mutate(change_aoi = c(0, cumsum(eez_group[-1L] != eez_group[-length(eez_group)] |
    as.numeric(difftime(date[-1L],
      date[-length(date)],
      units = "day"
    )) > 14))) %>%
  ungroup() %>%
  dplyr::select(ssvid, eez_group, date, change_aoi, everything()) %>%
  ungroup()

# find the start and end points of the change in aoi for segmenting 
time_range_data2 <- eez_act %>%
  arrange(date) %>%
  group_by(ssvid) %>%
  mutate(change_aoi = c(0, cumsum(eez_group[-1L] != eez_group[-length(eez_group)] |
    as.numeric(difftime(date[-1L],
      date[-length(date)],
      units = "day"
    )) > 14))) %>%
  ungroup() %>%
  dplyr::select(ssvid, eez_group, date, change_aoi, everything()) %>%
  group_by(ssvid, change_aoi) %>%
  summarise(
    first_date = min(date),
    last_date = max(date)
  ) %>%
  ungroup()

# join two datasets together 
merged_data2 <- base_data2 %>%
  inner_join(time_range_data2 %>%
    dplyr::select(ssvid, first_date, last_date),
  by = c("ssvid" = "ssvid", "date" = "first_date")
  ) %>%
  dplyr::select(ssvid, eez_group, date, last_date, everything()) %>%
  rename(first_date = date) %>%
  # rbind(aoi_padding) %>%
  mutate(aoi3 = factor(eez_group, levels = c(
    "nw_pacific","se_pacific", "sw_atlantic", "nw_indian",
    "other_eez","sw_indian_eezs","bahia_posesion",
    "nw_pacific_eezs","se_pacific_eezs","sw_atlantic_eezs"
  ))) %>%
  filter(!is.na(ssvid))

#-----------------------------------
# add eez and high seas dfs together 
# pull the merged dfs together 
combined <- rbind(dplyr::select(merged_data, ssvid, aoi3, first_date, last_date), 
      dplyr::select(merged_data2, ssvid, aoi3, first_date, last_date))

table(combined$aoi3)
# reset some levels for plotting 
# nw pacific and nw indian low levels so drop out  
combined$aoi3 <- as.character(combined$aoi3)
# combined$aoi3[combined$aoi3 %in% c('nw_pacific', 'nw_indian')] <- 'other_hs'
combined$aoi3[combined$aoi3 %in% c('nw_pacific', 'nw_indian')] <- NA
combined$ssvid <- factor(combined$ssvid)

# need to establish plotting order - currently desceding order of ssvid
# change to group by operating area to distinguish chinese returning vessels 

# gives the most common AOI to sort on 
order_data <- combined %>%
  arrange(first_date) %>%
  group_by(ssvid, aoi3) %>%
  mutate(
    duration = as.numeric(difftime(last_date, first_date , tz = "UTC", "secs")),
    duration = ifelse(first_date == last_date, 86400, duration)
  ) %>%
  ungroup() %>%
  arrange(first_date) %>%
  group_by(ssvid, aoi3) %>%
  summarize(total_days = sum(duration)) %>%
  ungroup() %>%
  group_by(ssvid) %>%
  filter(total_days == max(total_days)) %>%
  select(ssvid, most_common_aoi = aoi3) %>%
  inner_join(merged_data,
    by = c("ssvid" = "ssvid")
  )

# want to group vessels which return to China together - ID them first  
aoi_time <- combined %>%
  arrange(first_date) %>%
  group_by(ssvid, aoi3) %>%
  mutate(
    duration = as.numeric(difftime(last_date, first_date , tz = "UTC", "secs")),
    duration = ifelse(first_date == last_date, 86400, duration)
  ) %>%
  ungroup() %>%
  arrange(first_date) %>%
  group_by(ssvid, aoi3) %>%
  summarize(total_days = sum(duration)) %>%
  ungroup() %>%
  group_by(ssvid) 

# # vessels with > 1 month in NW Pacific EEZ aka China  
# chn_return <- aoi_time %>% 
#                 filter(aoi3=='nw_pacific_eezs') %>% 
#                 # time currently in seconds so multiply up to days
#                  filter(total_days >= (60*60*24*90))

# make the focus on going to Bahia Posesion rather than return to CHN
bp_test <- combined %>% 
  mutate(year= year(first_date)) %>% 
  group_by(ssvid, year) %>% 
  summarise(return = ifelse(aoi3 == 'bahia_posesion', TRUE, FALSE)) %>% 
  distinct() %>% 
  filter(return)
  
return_both <- bp_test %>% 
                  ungroup() %>% 
                  group_by(ssvid) %>% 
                  summarise(n = n()) %>% 
                  filter(n>1)
return_2021 <- filter(bp_test, year == '2021' & !ssvid %in% return_both$ssvid)
return_2022 <- filter(bp_test, year == '2022' & !ssvid %in% return_both$ssvid)

# correct most common aoi of the chn return vessels
order_data$most_common_aoi[order_data$ssvid %in% return_2021$ssvid] <- 'b_return_21'
order_data$most_common_aoi[order_data$ssvid %in% return_2022$ssvid] <- 'a_return_22'
order_data$most_common_aoi[order_data$ssvid %in% return_both$ssvid] <- 'c_return_bth'
table(order_data$most_common_aoi)


# create p_order column to try force plotting order. 
order_data$p_order <- paste(order_data$most_common_aoi, order_data$ssvid, sep='_')


# make segments with p_order as a factor to dictate the plotting order. 
segment_data <- order_data %>%
  mutate(
      ssvid_order = paste0(most_common_aoi, ssvid)
    ) %>%
  group_by(p_order, ssvid) %>%
  summarize(
    start = min(first_date, na.rm = TRUE),
    end = max(last_date, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(ssvid2 = factor(p_order))

# replicate factor level within the combined dataset
test <- combined %>% 
          left_join(dplyr::select(segment_data, ssvid, ssvid2)) %>% 
          arrange(desc(ssvid2))

# examine if plotting df appears as expected
#View(test %>% arrange(ssvid) %>% filter(grepl("prc", ssvid2, fixed = TRUE)))


p <- ggplot() +
  geom_segment(
    data = segment_data,
    aes(
      x = start,
      xend = end,
      # y = ssvid,
      # yend = ssvid,      
      y = ssvid2,
      yend = ssvid2,
    ),
    color = "grey90"
  ) +
  geom_segment(
    data = test,
    # data = combined,
    aes(
      x = first_date,
      xend = last_date,
      # y = ssvid,
      # yend = ssvid,
      y = ssvid2,
      yend = ssvid2,
      color = aoi3
    ),
    alpha = 1
  ) +
  scale_color_manual("Fishing Region",
    breaks = c(
    #"nw_pacific", "sw_atlantic", "nw_indian",
    "sw_atlantic","sw_atlantic_eezs",
    "other_eez","sw_indian_eezs","bahia_posesion",
    "nw_pacific_eezs","se_pacific", "se_pacific_eezs"
    ),
    # values = c("#5D69B1", "#58E8C6", "#99C945", "#CC3A8E",
    #            "#465192", "#148f72","#587622","#822259"),    
    # values = c("#00C1E7","#F59E84","#CC3A8E","#99C945","#ACBABF","#24796C", "#000000", # "#764E9F",
    #            "#CA7400","#5D69B1","#BE2045","#58E8C6","#A5AA99"),    
    values = c("#3EF7FF","#4AAACE","#ACBABF","#24796C", "#000000", "#CA7400","#FF6386","#C05282"),
    # labels = c(
    #   "NW Pacific Ocean High Seas", "SE Pacific Ocean High Seas",
    #   "SW Atlantic Ocean High Seas", "NW Indian Ocean High Seas", "Other EEZ",
    #   "SW Indian Ocean EEZs","Bahia Posesion", "NW Pacific Ocean EEZs", "SE Pacific Ocean EEZs",
    #   "SW Atlantic Ocean EEZs", "NW Indian Ocean EEZs"   
      labels = c(
      "SW Atlantic Ocean High Seas", "SW Atlantic Ocean EEZs", "Other EEZ",
      "SW Indian Ocean EEZs","Bahia Posesion", "NW Pacific Ocean EEZs", 
      "SE Pacific Ocean High Seas","SE Pacific Ocean EEZs"
    )
  ) +
  scale_x_date("",
    date_minor_breaks = "months",
    date_breaks = "year",
    date_labels = "%Y",
    limits = c(as.Date("2021-01-01"), as.Date("2022-12-31")),
    expand = c(0, 0)
  ) +
  theme_gfw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    # axis.text.y = element_text(size = 1),
    axis.text.y = element_blank(),
    panel.border = element_rect(color = "grey85", fill = NA),
    panel.background = element_rect(fill = "#ffffff"),
    plot.background = element_rect(fill = "#ffffff"),
    strip.text.y.left = element_text(angle = 0),
    panel.grid.minor.x = element_line(color = "grey85", linetype = "dotted"),
    axis.title = element_blank(),
    legend.position = "bottom",
    legend.text=element_text(size=6),
    legend.title=element_text(size=7),
    plot.title = element_text(family = "Roboto-Regular")
  ) 
  p
getwd()
ggsave('./outputs/figures/Bahia_posesion_draft_linkage_ordered2.png',plot=p, device='png')
ggsave('../outputs/figures/Bahia_posesion_draft_linkage_ordered.png',plot=p, device='png')


```


```{r}
bp_enc_q <- readr::read_file(file = here::here("queries", ".", "Bahia_Posesion_encounters.sql"))

# bp_enc <- fishwatchr::gfw_query(query = glue::glue(bp_enc_q,
#                                               date_start = '2021-01-01',
#                                               date_end = '2022-11-30'),
bp_enc <- fishwatchr::gfw_query(query = bp_enc_q,
                                 run_query = TRUE,
                                 con = con)$data

str(bp_enc)

# duplicate rows of ecounters 
unq_enc <- bp_enc %>% filter(!duplicated(event_start, event_end)) 

getwd()
write.csv(dplyr::select(unq_enc, -eez, -rfmo), '../data/Bahia_Posesion_encounters.csv')
```

```{r}

bp_q <- c("
SELECT 
  *
FROM `world-fishing-827.scratch_max.bahia_posesion_raw_ais`
")


bp_ais <- fishwatchr::gfw_query(query = bp_q,
                                 run_query = TRUE, 
                                 con = con)$data


ggplot(bp_ais) + geom_boxplot(aes(x=speed))


write.csv(bp_q,'data/BP_raw_ais.csv')

```


```{r, include=FALSE, warning=FALSE}
# try dockering back end function components to make a bespoke plot 

# plot time series component to show the vessels speed through time  
plot_df <- voi_atl %>% filter(between(as.Date(timestamp),as.Date('2022-01-18'),
                                                 as.Date('2022-01-24'))) 

# make the time series data in the correct format
ts_plot_df <- plot_df %>%
  dplyr::mutate(next_time = dplyr::lead(timestamp),
                time_diff_hr = as.numeric(difftime(next_time,
                                                   timestamp,
                                                   tz = 'UTC',
                                                   units = 'sec'))/3600,
                next_far = ifelse(time_diff_hr > 3, TRUE, FALSE),
                prev_far = ifelse(lead(next_far) == TRUE, TRUE, FALSE),
                nnet_score = as.character(ifelse(next_far == TRUE | prev_far == TRUE | 
                                                   is.na(nnet_score), 2 , nnet_score)),
                slow_speed = ifelse(speed_knots<5, TRUE, FALSE))

# make the line plot coloured by speed
line_geom <-
  geom_path(data = ts_plot_df,
            aes(x= timestamp,
                #eval(as.symbol(speed_knots)),
                y= speed_knots,
                color = factor(slow_speed),
                size = factor(slow_speed),
                group  = ssvid))
# set colours 
line_colorscale <-
  scale_color_manual('',
                     breaks = c(FALSE,TRUE),
                     values = c('#76ADAD',
                                '#E34D4E',
                                'grey85'),
                     # labels = NULL)
                     labels = c('>5 knots','< 5 knots'))
# set size 
line_sizescale <- scale_size_manual('',
                  breaks = c(TRUE,FALSE),
                  values = c(0.5,0.5,0.3),
                  # labels = NULL)
                  labels = c('>5 knots','< 5 knots'))

# time series plot of speed
time_series_plot <- ggplot() +
  line_geom +
  # point_geom +
  line_sizescale +
  line_colorscale +
  #y_axis +
  scale_x_datetime('\nDate') +
  fishwatchr::theme_gfw() +
  theme(
        legend.position = 'none',
        legend.justification = 'right',
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_rect(fill = '#ffffff'),
        panel.border = element_rect(fill = NA, color = '#000000'),
        plot.background = element_rect(fill = '#ffffff'))+
        # axis.text.x = element_blank(),
        # axis.title.x = element_blank()) +
  labs(y='Speed (knots)')

# make land object at default scale of  10 
land_sf <- rnaturalearth::ne_countries(scale = 10,
                                       returnclass = "sf")

# manual bounding box bounds set to +/- buf 
buf <- 0.1
bounding_box <- transform_box(xlim = c(min(ts_plot_df$lon) - buf,
                                               max(ts_plot_df$lon) + buf),
                                      ylim = c(min(ts_plot_df$lat) - buf,
                                               max(ts_plot_df$lat) + buf),
                                      output_crs = gfw_projections('Equal Earth')$proj)

# make a track object colored by speed 
track_sf <- make_track(ts_plot_df,
                       color_by = 'slow_speed')

# pull out vessel mmsi 
vessel_mmsi <- track_sf %>%
    sf::st_drop_geometry(.) %>%
    dplyr::distinct(ssvid) %>%
    dplyr::pull(ssvid)

# set land to light theme colours 
land_fill = fishwatchr::gfw_palettes$map_country_light[1]
land_color = fishwatchr::gfw_palettes$map_country_light[2]

# set grid colour 
grid_color = 'grey85'

# set vessel track geom based on speed
vessel_track_geom <-
  geom_sf(data = track_sf,
          aes(
            color = slow_speed,
            group = ssvid
          )
      )

# set geom color scale to speed too. 
vessel_track_colorscale <-
  scale_color_manual("",
                     breaks = c(FALSE, TRUE),
                     values = c(
                       "#76ADAD",
                       "#E34D4E",
                       'grey85'
                     ),
                     labels = c(
                       "> 5 knots",
                       "< 5 knots")
                     )

vessel_name <- 'LU RONG YUAN YU 715'

map <- ggplot() +
  geom_sf(
    data = land_sf,
    fill = land_fill,
    color = land_color,
    size = 0.2
  ) +
  vessel_track_geom +
  vessel_track_colorscale +
    geom_gfw_eez(colour='black', lty=2,lwd=1.2, theme ='light', proj = "+proj=eqearth +datum=WGS84 +wktext") +
  coord_sf(
    xlim = c(bounding_box$box_out[["xmin"]], bounding_box$box_out[["xmax"]]),
    ylim = c(bounding_box$box_out[["ymin"]], bounding_box$box_out[["ymax"]]),
    crs = bounding_box$out_crs
  ) +
  fishwatchr::theme_gfw_map(theme = 'light') +
  labs(title = vessel_name,
       subtitle = vessel_mmsi, 
       fill = 'Speed') +
    theme(
    # legend.position = c(1.7,0.2),
    legend.position = 'bottom',
    # legend.position = 'none',
    legend.justification = "right",
    # panel.border = element_rect(fill = NA, color = 'grey15'),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    panel.grid.major = element_line(colour = grid_color,
                                    size = 0.1), 
    plot.title=element_text(size=10),
    plot.subtitle=element_text(size=9)
  ) 

# stacked_timeseries <- stack_timeseries(line_plot_list)
# num_components <- length(line_plot_list)
# height_ratio_list <- list(0.2, 0.3, 0.35, 0.4, 0.5)

num_components <- 1
page_layout <- patchwork::plot_layout(ncol = 1,
                                      # widths = c(1.5,1),
                                      heights = c(1 - 0.3, 0.3),
                                      guides = 'auto')
                                      # guides = 'keep')

# arrange plots on page
fishwatchr::add_little_globe(map,
                             main_box = bounding_box,
                             globe_rel_size = '0.25',
                             globe_just = 'inside',
                             globe_position = 'upperright') +
    time_series_plot +
    page_layout

# map +
#     time_series_plot +
#     page_layout

# setwd('./outputs/figures')
ggsave('../outputs/figures/LU RONG YUAN YU 715_example_Atlantic.png', device='png')



```

Show example image of vessel activity in Bahia Posesion. 



```{r, include=FALSE, warning=FALSE}
voi_q <- c("
           SELECT 
              *
            -- FROM `world-fishing-827.pipe_production_v20201001.messages_segmented_20221111`
            -- FROM `world-fishing-827.pipe_production_v20201001.messages_scored_20221111`
            -- FROM `pipe_ais_sources_v20201001.normalized_orbcomm_*`
            FROM `pipe_ais_sources_v20201001.normalized_spire_*`
            WHERE   
            ssvid = '{voi}' AND 
            DATE(timestamp) BETWEEN DATE('{start_date}') AND DATE('{end_date}')
           ")

voi_ais_segs <- fishwatchr::gfw_query(query = glue::glue(voi_q,
                                               voi = '412440777', 
                                               start_date = '2021-05-05',
                                               end_date = '2021-05-30'),
                                 run_query = TRUE,
                                 con = con)$data
nrow(voi_ais_segs)
voi_ais_segs <- voi_ais_segs %>% filter(!is.na(lon), !is.na(lat)) %>% arrange(timestamp)
nrow(voi_ais_segs)

```

```{r, include=FALSE, warning=FALSE}
# use bahia posesion ais as starting point for aoi 

voi <- filter(bp_ais, ssvid == 412440777)
min(voi$timestamp)
max(voi$timestamp)


# make the time series data in the correct format
ts_plot_df <- voi %>%
  dplyr::mutate(next_time = dplyr::lead(timestamp),
                time_diff_hr = as.numeric(difftime(next_time,
                                                   timestamp,
                                                   tz = 'UTC',
                                                   units = 'sec'))/3600,
                next_far = ifelse(time_diff_hr > 3, TRUE, FALSE),
                prev_far = ifelse(lead(next_far) == TRUE, TRUE, FALSE),
                slow_speed = ifelse(speed<5, TRUE, FALSE))

# manual bounding box bounds set to +/- buf 
buf <- 0.1
bounding_box <- transform_box(xlim = c(min(ts_plot_df$lon) - buf,
                                               max(ts_plot_df$lon) + buf),
                                      ylim = c(min(ts_plot_df$lat) - buf,
                                               max(ts_plot_df$lat) + buf),
                                      output_crs = gfw_projections('Equal Earth')$proj)

# use longer time series raw data to show larger movement into and from BP 

ts_plot_df <- dplyr::select(voi_ais_segs, ssvid, timestamp, lon, lat, speed) %>%
  dplyr::mutate(next_time = dplyr::lead(timestamp),
                time_diff_hr = as.numeric(difftime(next_time,
                                                   timestamp,
                                                   tz = 'UTC',
                                                   units = 'sec'))/3600,
                next_far = ifelse(time_diff_hr > 3, TRUE, FALSE),
                prev_far = ifelse(lead(next_far) == TRUE, TRUE, FALSE),
                slow_speed = ifelse(speed<5, TRUE, FALSE))



# make the line plot coloured by speed
line_geom <-
  geom_path(data = ts_plot_df,
            aes(x= timestamp,
                #eval(as.symbol(speed_knots)),
                y= speed,
                color = factor(slow_speed),
                size = factor(slow_speed),
                group  = ssvid))
# set colours 
line_colorscale <-
  scale_color_manual('',
                     breaks = c(FALSE,TRUE),
                     values = c('#76ADAD',
                                '#E34D4E',
                                'grey85'),
                     # labels = NULL)
                     labels = c('>5 knots','< 5 knots'))
# set size 
line_sizescale <- scale_size_manual('',
                  breaks = c(TRUE,FALSE),
                  values = c(0.5,0.5,0.3),
                  # labels = NULL)
                  labels = c('>5 knots','< 5 knots'))

# time series plot of speed
time_series_plot <- ggplot() +
  line_geom +
  # point_geom +
  line_sizescale +
  line_colorscale +
  #y_axis +
  scale_x_datetime('\nDate') +
  fishwatchr::theme_gfw() +
  theme(
        legend.position = 'none',
        legend.justification = 'right',
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_rect(fill = '#ffffff'),
        panel.border = element_rect(fill = NA, color = '#000000'),
        plot.background = element_rect(fill = '#ffffff'))+
        # axis.text.x = element_blank(),
        # axis.title.x = element_blank()) +
  labs(y='Speed (knots)')

# make land object at default scale of  10 
land_sf <- rnaturalearth::ne_countries(scale = 10,
                                       returnclass = "sf")


# make a track object colored by speed 
track_sf <- make_track(ts_plot_df,
                       color_by = 'slow_speed')

# pull out vessel mmsi 
vessel_mmsi <- track_sf %>%
    sf::st_drop_geometry(.) %>%
    dplyr::distinct(ssvid) %>%
    dplyr::pull(ssvid)

# set land to light theme colours 
land_fill = fishwatchr::gfw_palettes$map_country_light[1]
land_color = fishwatchr::gfw_palettes$map_country_light[2]

# set grid colour 
grid_color = 'grey85'

# set vessel track geom based on speed
vessel_track_geom <-
  geom_sf(data = track_sf,
          aes(
            color = slow_speed,
            group = ssvid
          )
      )

# set geom color scale to speed too. 
vessel_track_colorscale <-
  scale_color_manual("",
                     breaks = c(FALSE, TRUE),
                     values = c(
                       "#76ADAD",
                       "#E34D4E",
                       'grey85'
                     ),
                     labels = c(
                       "> 5 knots",
                       "< 5 knots")
                     )

vessel_name <- 'FU YUAN YU 7628'

map <- ggplot() +
  geom_sf(
    data = land_sf,
    fill = land_fill,
    color = land_color,
    size = 0.2
  ) +
  vessel_track_geom +
  vessel_track_colorscale +
    geom_gfw_eez(colour='black', lty=2,lwd=1.2, theme ='light', proj = "+proj=eqearth +datum=WGS84 +wktext") +
  coord_sf(
    xlim = c(bounding_box$box_out[["xmin"]], bounding_box$box_out[["xmax"]]),
    ylim = c(bounding_box$box_out[["ymin"]], bounding_box$box_out[["ymax"]]),
    crs = bounding_box$out_crs
  ) +
  geom_rect(aes(xmin = -69.40,
                xmax = -69.13,
                ymin = -52.31,
                ymax = -52.20), color = "red", fill = NA)  +
  fishwatchr::theme_gfw_map(theme = 'light') +
  labs(title = vessel_name,
       subtitle = vessel_mmsi, 
       fill = 'Speed') +
    theme(
    # legend.position = c(1.7,0.2),
    legend.position = 'bottom',
    # legend.position = 'none',
    legend.justification = "right",
    # panel.border = element_rect(fill = NA, color = 'grey15'),
    plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
    panel.grid.major = element_line(colour = grid_color,
                                    size = 0.1), 
    plot.title=element_text(size=10),
    plot.subtitle=element_text(size=9)
  ) 

# stacked_timeseries <- stack_timeseries(line_plot_list)
# num_components <- length(line_plot_list)
# height_ratio_list <- list(0.2, 0.3, 0.35, 0.4, 0.5)

num_components <- 1
page_layout <- patchwork::plot_layout(ncol = 1,
                                      # widths = c(1.5,1),
                                      heights = c(1 - 0.3, 0.3),
                                      guides = 'auto')
                                      # guides = 'keep')

# arrange plots on page
fishwatchr::add_little_globe(map,
                             main_box = bounding_box,
                             globe_rel_size = '0.25',
                             globe_just = 'inside',
                             globe_position = 'upperright')+ 
    time_series_plot +
    page_layout
# 
# getwd()

# setwd('./outputs/figures')
ggsave('../outputs/figures/FU YUAN YU 7628_example_BahiaPosesion_withTS.png', device='png')
```
```
